---
title: Network Model
---

Turns out I am not very good at.

## Spatial Analyis

### Recommendation 

- For visualize large data set use: `rdeck`
- `sf` is standard for spatial transformation
- `spatstat::as.ppp()` spatial point data for computation. 


```{r}
library(sfnetworks)
library(sf)
library(terra)
library(stars)
library(spatstat)
#remotes::install_github("qfes/rdeck")
library(rdeck) # lightweight spatial map
```


```{r}
#install.packages('spatstat')
#ll= st_as_sf(longleaf)
```
```{r}
#lll |> as.ppp()
```
### Observation Window
```{r}
n <- 30
set.seed(13531) # remove this to create another random sequence
xy <- data.frame(x = runif(n), y = runif(n)) |> 
    st_as_sf(coords = c("x", "y"))
w1 <- st_bbox(c(xmin = 0, ymin = 0, xmax = 1, ymax = 1)) |> 
        st_as_sfc() 
w2 <- st_sfc(st_point(c(1, 0.5))) |> st_buffer(1.2)
par(mfrow = c(1, 2), mar = c(2.1, 2.1, 0.1, 0.5), xaxs = "i", yaxs = "i")
plot(w1, axes = TRUE, col = 'grey')
plot(xy, add = TRUE)
plot(w2, axes = TRUE, col = 'grey')
plot(xy, add = TRUE, cex = .5)
```

use `as.ppp` function
```{r}
(pp1 <- c(w1, st_geometry(xy)) |> as.ppp())
# Planar point pattern: 30 points
# window: polygonal boundary
# enclosing rectangle: [0, 1] x [0, 1] units
c1 <- st_buffer(st_centroid(w2), 1.2)
(pp2 <- c(c1, st_geometry(xy)) |> as.ppp())
# Planar point pattern: 30 points
# window: polygonal boundary
# enclosing rectangle: [-0.2, 2.2] x [-0.7, 1.7] units
par(mfrow = c(1, 2), mar = rep(0, 4))
q1 <- quadratcount(pp1, nx=3, ny=3)
q2 <- quadratcount(pp2, nx=3, ny=3)
plot(q1, main = "")
plot(xy, add = TRUE)
plot(q2, main = "")
plot(xy, add = TRUE)
```



```{r}
den1 <- density(pp1, sigma = bw.diggle)
den2 <- density(pp2, sigma = bw.diggle)
par(mfrow = c(1, 2), mar = c(0,0,1.1,2))
plot(den1, col = colorRampPalette(RColorBrewer::brewer.pal(9,"BuGn"))(100))
plot(pp1, add=TRUE)
plot(den2)
plot(pp1, add=TRUE)
```
Here is code to replicate this plot
Adjust aesthetics to fit my demand. 
```{r}

xy <- data.frame(x = runif(n), y = runif(n)) |> 
    st_as_sf(coords = c("x", "y"))

plotPPPdensity = function(xy) {
  maxRange =xy |> sf::st_coordinates() |> apply(2, max)
  minRange =xy |> sf::st_coordinates() |> apply(2, min)
  w1 <- sf::st_bbox(c(xmin = minRange[['X']], 
                  ymin = minRange[['Y']], 
                  xmax = maxRange[['X']], 
                  ymax = maxRange[['Y']])) |> 
          sf::st_as_sfc()
  # xy |> st_combine() |> st_centroid()
  # w1 |> st_centroid()
  #   st_buffer(0.2)
  (pp1 <- c(w1, sf::st_geometry(xy)) |> spatstat.geom::as.ppp())
  den1 <- density.ppp(pp1, sigma = bw.diggle)
  par(mar =c(0.2, 0.2, 0.2, 0.2))
  plot(den1, 
       col = colorRampPalette(RColorBrewer::brewer.pal(9,"BuGn"))(500),
       main=NULL,
       legend=F, show.all=F
       )
  plot(pp1, add = T)
}
xy2 <- data.frame(x = rnorm(50, 1, 0.3), y = rnorm(50, 1, 0.3)) |> 
    st_as_sf(coords = c("x", "y"))
plotPPPdensity(xy2)
```

```{r}

```


```{r}
xy |> class()
g = igraph::make_tree(30, 5)
V(g)$geom = xy$geometry
class(V(g)$geom)
class(V(g)$geom[[1]])
V(g)$geom |> st_as_sf() |> plotPPPdensity()

library(shiny)
ui <- fluidPage(
  fluidRow(plotOutput("id", brush="plot_brush"), click="plot_click"),
  fluidRow(verbatimTextOutput("dev"))
)

server <- function(input, output, session) {
  output$id <- renderPlot({
    plotPPPdensity(xy)
  })
  observe({
    
  })
  output$dev <- renderPrint({
    print(input$plot_click)
    print(input$plot_brush)
  })
}
shinyApp(ui, server)
```

These are the output by selecting just one point... 
```
$xmin
[1] 0.2553337

$xmax
[1] 0.5582186

$ymin
[1] 0.175992

$ymax
[1] 0.4206298
```
I need to figure out if those scales are in absolute or relative value

```{r}
xy2
w = st_bbox(c())
```

```{r}
library(shiny)
library(ggplot2)
# graph definition
g = make_tree(50, 5)
V(g)$geom <- xy2$geometry
    
ui <- fluidPage(
  fluidRow(
    sidebarPanel(
      div(plotOutput("id", 
                 brush="plot_brush", 
                 click="plot_click",
                 height="300px", width="300px"
                 ))
    ),
    mainPanel(
      visNetwork::visNetworkOutput("visNet")
    )
  ),
  fluidRow(verbatimTextOutput("dev"))
)

server <- function(input, output, session) {
  output$id <- renderPlot({
    plotPPPdensity({
      V(g)$geom |> st_as_sf()
    })
  })
  output$visNet <- visNetwork::renderVisNetwork({
    visNetwork::visIgraph(g) |> 
      visNetwork::visOptions(highlightNearest = list(
        enabled=T, degree=0
      ))
  })
  observe({
    req(cur)
    cur = input$plot_brush
    brash_area = st_bbox(c(xmin =cur$xmin, xmax = cur$xmax, ymin=cur$ymin, ymax =cur$ymax)) |> 
      st_as_sfc()
    selected_dots =st_intersects(st_as_sf(V(g)$geom), brash_area, sparse=F) |> which()
    visNetwork::visNetworkProxy("visNet") |> 
      visNetwork::visSelectNodes(selected_dots)
  })
  output$dev <- renderPrint({
    # print(input$plot_click)
    # print(input$plot_brush)
    cur = input$plot_brush
    req(cur)
    x = st_bbox(c(xmin =cur$xmin, xmax = cur$xmax, ymin=cur$ymin, ymax =cur$ymax)) |> 
      st_as_sfc()
    c =st_intersects(xy2, x, sparse=F)
    print(c)
    print(which(c))
  })
}
shinyApp(ui, server)
```

### Continute 
Previously: 

1. sf::pointgeometry can be converted to startstat::ppp 
2. convert this startstat::ppp -> image via method `density`
3. **it is possible to extract value from image via st_as_stars**


```{r}
xy_density = xy |> as.ppp() |> density() |> st_as_stars()
xy_density$area = xy_density |> st_area()

with(xy_density, sum(v * area, na.rm=T))
```
```{r}
pt <- st_sfc(st_point(c(0.5, 0.5)))
st_as_sf(xy_density, as_points = TRUE, na.rm = FALSE) |>
  st_distance(pt) -> xy_density$dist

(m <- ppm(pp2 ~ dist, data = list(dist = as.im(xy_density["dist"]))))

plot(m)
```

